Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-01 paper/paper.tex
new Cortex
Change RL/POMDP figure
Toned down UK Biobank claims

Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
the brain is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
The present paper proposes a
process model that satisfies this previously proposed contention.
We also contribute to the discussion of DMN function
by providing some of the first empirical evidence that
morphological variability in DMN regions
is linked to the reward circuitry (Fig. 2),
thus linking two literatures with currently scarce cross-references.
Finally, we detail how our process model relates
to previous accounts of DMN function and
we derive explicit hypotheses to be tested in future neuroscience
experiments.
At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
%
Notably, our account does not claim that the brain or neural activity in the DMN are completely identical with reinforcement learning algorithms. Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
to understand DMN function.

@@

% \begin{wrapfigure}{r}{.4\linewidth}
\begin{figure}[!htbp]
  \begin{minipage}[c]{0.63\textwidth}
    \includegraphics[width=.8\linewidth]{fig_dmn.png}
  \end{minipage}
  \hspace{-4em}
  \begin{minipage}[c]{0.45\textwidth}
    \caption{\textbf{Morphological coupling between reward system and
        default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN regions. These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
      Consistent for the left and right reward system, NAc volume in a given subject is
      positively coupled with the vmPFC and HC. The colors are indicative of the (\textcolor{red}{red} = positive,
      \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  The code for reproduction and visualization:
  \url{www.github.com/banilo/darkcontrol_PCB2018}.
  }
  \label{fig:dmn}
\end{minipage}
% \end{wrapfigure}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=1.\linewidth]{rl2.pdf}
  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
    the agent takes an action by following the policy matrix as updated by the Bellman equation. The agent receives a triggered reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
  \label{fig:rl}
\end{figure}

>>>>

Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
the brain is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
The present paper proposes a
process model that satisfies this previously proposed contention.
We also contribute to the discussion of DMN function
by providing tentative evidence that
variation of the grey-matter volume in DMN regions
is linked to the reward circuitry (Fig. 2),
thus linking two literatures with currently scarce cross-references.
Finally, we detail how our process model relates
to previous accounts of DMN function and
we derive explicit hypotheses to be tested in future neuroscience
experiments.
At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
%
Notably, our account does not claim that the brain or neural activity in the DMN are completely identical with reinforcement learning algorithms. Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
to understand DMN function.

@@

% \begin{wrapfigure}{r}{.4\linewidth}
\begin{figure}[!htbp]
  \begin{minipage}[c]{0.63\textwidth}
    \includegraphics[width=.8\linewidth]{fig_dmn.png}
  \end{minipage}
  \hspace{-4em}
  \begin{minipage}[c]{0.45\textwidth}
    \caption{\textbf{Morphological coupling between reward system and
        default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN regions. These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
      Consistent for the left and right reward system, NAc volume in a given subject is
      positively coupled with the vmPFC and HC. The colors are indicative of the (\textcolor{red}{red} = positive,
      \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  The code for reproduction and visualization:
  \url{www.github.com/banilo/darkcontrol_2018}.
  }
  \label{fig:dmn}
\end{minipage}
% \end{wrapfigure}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=1.\linewidth]{POMDP_Fig.pdf}
  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
    the agent takes an action by following the policy matrix as updated by the Bellman equation. The agent receives a triggered reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
  \label{fig:rl}
\end{figure}
